Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system with plausible mechanism

### 2024-07-17 _tex/index.tex
Built site for gh-pages

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Causes and effects in Dichotomous Comparative Judgments: a plausible information-theoretical system of mechanism},
  pdfauthor={Jose Manuel Rivera Espejo; Tine van van Daal; Sven De De Maeyer; Steven Gillis},
  pdfkeywords={comparative judgement, directed acycilc graph, causal
analysis, probabilistic statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Causes and effects in Dichotomous Comparative Judgments: a
plausible information-theoretical system of mechanism}
\author[1]{Jose Manuel Rivera Espejo%
\corref{cor1}%
}
 \ead{JoseManuel.RiveraEspejo@uantwerpen.be} 
\author[1]{Tine van Daal%
%
}
 \ead{tine.vandaal@uantwerpen.be} 
\author[1]{Sven De Maeyer%
%
}
 \ead{sven.demaeyer@uantwerpen.be} 
\author[2]{Steven Gillis%
%
}
 \ead{steven.gillis@uantwerpen.be}

\affiliation[1]{organization={University of Antwerp, Education
sciences},,postcodesep={}}
\affiliation[2]{organization={University of
Antwerp, Linguistics},,postcodesep={}}

@@

@@

\textbf{Code availability:} All the code utilized in this research is
available in the different notebooks and \texttt{CODE\ LINKS} referenced
in the digital document. The digital document is located at:
\url{https://jriveraespejo.github.io/paper2_manuscript/}.

\textbf{Authors' contributions:} \emph{Conceptualization:} S.G., S.dM.,
and J.M.R.E; \emph{Data curation:} J.M.R.E.; \emph{Formal Analysis:}
J.M.R.E.; \emph{Funding acquisition:} S.G. and S.dM;
\emph{Investigation:} S.G.; \emph{Methodology:} S.G., S.dM., and
J.M.R.E; \emph{Project administration:} S.G. and S.dM.;
\emph{Resources:} S.G. and S.dM.; \emph{Software: J.M.R.E.};
\emph{Supervision:} S.G. and S.dM.; \emph{Validation:} J.M.R.E.;
\emph{Visualization:} J.M.R.E.; \emph{Writing - original draft:}
J.M.R.E.; \emph{Writing - review \& editing:} S.G. and S.dM.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system with plausible mechanism},
  pdfauthor={Jose Manuel Rivera Espejo; Tine van van Daal; Sven De De Maeyer; Steven Gillis},
  pdfkeywords={comparative judgement, directed acycilc graph, causal
analysis, probabilistic statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Causes and effects in Dichotomous Comparative Judgments: an
information-theoretical system with plausible mechanism}
\author[1]{Jose Manuel Rivera Espejo%
\corref{cor1}%
}
 \ead{JoseManuel.RiveraEspejo@uantwerpen.be} 
\author[1]{Tine van Daal%
%
}
 \ead{tine.vandaal@uantwerpen.be} 
\author[1]{Sven De Maeyer%
%
}
 \ead{sven.demaeyer@uantwerpen.be} 
\author[2]{Steven Gillis%
%
}
 \ead{steven.gillis@uantwerpen.be}

\affiliation[1]{organization={University of Antwerp, Training and
education sciences},,postcodesep={}}
\affiliation[2]{organization={University of
Antwerp, Linguistics},,postcodesep={}}

@@

In contemporary contexts, Thurstone's law of comparative judgment
\citeyearpar{Thurstone_1927} primarily refers to the method of
\emph{Dichotomous} Comparative Judgment
\citep[DCJ,][]{Pollitt_2012a, Pollitt_2012b}. In DCJ, a judge assesses
the relative manifestation of a \emph{trait} within a pair of stimuli.
This assessment results in a dichotomous value indicating which stimulus
possesses a higher degree of the trait. After different judges perform
multiple rounds of pairwise comparisons, an outcome vector is produced.
This vector is modeled using the Bradley-Terry-Luce model
\citep[BTL,][]{Bradley_et_al_1952, Luce_1959, Pollitt_2012a}, which
produces a latent variable corresponding to the trait of interest. This
latent variable is then used to rank the stimuli from lowest to highest
or to evaluate the influence of certain variables on the stimuli's
positions in the ranking.

DCJ has proven effective in assessing competencies and traits
predominantly within the educational realm, as demonstrated by the works
of \citet{Pollitt_2012b}, \citet{Bartholomew_et_al_2018},
\citet{vanDaal_et_al_2016}, \citet{Lesterhuis_2018}, and
\citet{Bartholomew_et_al_2020}. However, its application transcends
education, as exemplified by the work of \citet{Boonen_et_al_2020}. The
methodology has also evolved to include multiple, as opposed to pairwise
comparisons \citep{Luce_1959, Placket_1975}, and to accommodate
comparisons with ordinal outcomes \citep{Tutz_1986, Agresti_1992}.
Overall, research suggests that DCJ offers an alternative and efficient
approach to measurement and evaluation, characterized by its reliability
and validity \citep{Lesterhuis_2018, vanDaal_2020, Marshall_et_al_2020}.
Nevertheless, despite the method's widespread use, the literature lacks
a transparent depiction of the DCJ system and the plausible mechanisms
that give rise to DCJ data. Particularly, there is no detailed
explanation of how different assessment factors can potentially
influence the observed DCJ data.

According to \citet{Verhavert_et_al_2019} and \citet{vanDaal_2020},
several assessment factors interact and contribute to the reliability of
the DCJ method. These factors include the number and characteristics of
the stimuli, their \emph{proximity} in terms of the assessed trait, the
number of comparison per stimulus, and the pairing algorithm used.
Furthermore, since the method relies on judges' assessments, the number
and characteristics of judges, their \emph{discrimination} abilities,
and the number of comparisons per judge also play pivotal roles.
Moreover, when the stimuli represent sub-units of higher-levels units,
factors such as the number and characteristics of these units, along
with their \emph{proximity} in terms of the assessed trait, can
significantly influence the outcome. An example of this can be found in
\citet{vanDaal_et_al_2016}, where the authors assessed the \emph{skills
in academic writing} (trait) of Flemish university students, utilizing
multiple argumentative essays (stimuli, sub-units) originating from
various students (units).

@@

\textbf{Code availability:} All the code utilized in this research is
available in the digital document located at:
\url{https://jriveraespejo.github.io/paper2_manuscript/}.

\textbf{Authors' contributions:} \emph{Conceptualization:} S.G., S.DM.,
T.vD., and J.M.R.E; \emph{Methodology:} S.DM., T.vD., and J.M.R.E;
\emph{Software:} J.M.R.E.; \emph{Validation:} J.M.R.E.; \emph{Formal
Analysis:} J.M.R.E.; \emph{Investigation:} J.M.R.E; \emph{Resources:}
S.G., S.DM., and T.vD.; \emph{Data curation:} J.M.R.E.; \emph{Writing -
original draft:} J.M.R.E.; \emph{Writing - review \& editing:} S.G.,
S.DM., and T.vD.; \emph{Visualization:} J.M.R.E.; \emph{Supervision:}
S.G. and S.DM.; \emph{Project administration:} S.G. and S.DM.;
\emph{Funding acquisition:} S.G. and S.DM.
